---
title: 'Text Regular Expressions'
author: "Illarion  Jabine"
date: "09/04/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


### 1. Required packages:

 * strinr: Simple, Consistent Wrappers for Common String Operations

### 2. Cheatsheet

1. Character classes
  .	any character except newline
  \w\d\s	word, digit, whitespace
  \W\D\S	not word, digit, whitespace
  [abc]	any of a, b, or c
  [^abc]	not a, b, or c
  [a-g]	character between a & g

2. Anchors 
  ^abc$	start / end of the string
  \\b \\B	word, not-word boundary

3. Escape characters
 \\. \\* \\	escaped special characters
 \\t \\n \\r	tab, linefeed, carriage return

4. Groups and Lookaround
 (abc)	capture group
 \1	backreference to group #1
 (?:abc)	non-capturing group
 (?=abc)	positive lookahead
 (?!abc)	negative lookahead

5. Quantifiers and alternation
 a* a+ a?	0 or more, 1 or more, 0 or 1
 a{5} a{2,}	exactly five, two or more
 a{1,3}	between one & three
 a+? a{2,}?	match as few as possible
 ab|cd	match ab or cd

### 3. Useful Links
 $ <https://r4ds.had.co.nz/strings.html>
 $ <https://stringr.tidyverse.org/articles/regular-expressions.html>
 
## 4. Introduction
Regular expressions are widely used in manipulating (parsing/replacing strings) text and extracting information from it.

### 5. Load the libraries
Let's first load the libraries.
```{r loading packages, message=FALSE, warning=FALSE}
library(stringr)
```

### 6. Loading and checking the data

Load and check the datasets.
Dataset1: movie_titles

```{r load the data and pre-process them}
# Loading data from Rds file
movie_titles <- c("Karate Kid","The Twilight Saga: Eclispe","Knight & Day","Shrek Forever After 3D","Marmaduke.","Street Dance","Predators","StreetDance 3D","Robin Hood","Micmacs A Tire-Larigot","50 Shades of Grey","Sex And the City 2","Inception","The Dark Knight","300","Toy Story 3 In Disney Digital 3D","50 Shades of Gray","Italien, Le","Tournee","The A-Team","El Secreto De Sus Ojos","Kiss & Kill","The Road","Cosa Voglio Di Piu","Nur für dich","Prince Of Persia: The Sands Of Time","Saw 4","Saw 5","Saw 6","21 Grams")
```

### 7. Special characters

Special character --> Meaning
^ --> Caret: Marks the beginning of a line or string ("starts with")
$ --> Dollar Sign: Marks the end of a line or string ("ends with")
. --> Period: Matches anything: letters, numbers or white spaces
\\. --> Two backslashes: Escapes the period when we search an actual period

Control characters:
 * \n: line feed
 * \r: carriage return 
 * \t: horizontal tabulation

```{r}
# Will match first two chars at the beginning. When creating a pattern to look for something at the beginning of a line, use the caret followed by a search term.
string <- c("This is a text string.","This is another text string","Some text?")
str_match(string, "^..")

# The period matches any character, it's like a wild card. Example, using three dots "..." will find three characters - be it letters or numbers or white spaces.


# Match all sequels of the movie "Saw"
str_match(movie_titles, pattern = "Saw..")

# Will match strings with "T" at the beginning
str_match(string, "^T")

# str_view_all() shows all the matches
str_view_all(movie_titles,"^The")

# Will match strings with "." at the end. When looking for something at the end, type the search term first and then append the dollar sign.
str_match(string, "\\.$")

# str_detect() returns TRUE if the pattern matches and FALSE otherwise, good for indexing.
str_detect("This is text","t$")

# List all movies that end with "3D"
movie_titles[str_detect(movie_titles,pattern = "3D$")]

```


### 8. Character Classes

 * \\d matches any digit: 0, 1, 2, 3,...
   The complement, \D, matches any character that is not a decimal digit.

 * \\w matches any “word” character, which includes alphabetic characters, marks and decimal numbers a, b, c, 1, 2, 3,...
 The complement, \W, matches any non-word character.
 
 * \s: matches any whitespace. This includes tabs, newlines, form feeds
 The complement, \\S, matches any non-whitespace character.

Your own character classes can be created using []:

[A-Za-z]: matches every character between A and Z and between a and z.
[abc]: matches a, b, or c.
[^abc]: matches anything except a, b, or c. A caret ^ inside the square brackets [^...] will invert the selection inside the brackets.

There are a number of pre-built classes that you can use inside []:
[:punct:]: punctuation.
[:alpha:]: letters.
[:lower:]: lowercase letters.
[:upper:]: upperclass letters.
[:digit:]: digits.
[:xdigit:]: hex digits.
[:alnum:]: letters and numbers.
[:cntrl:]: control characters.
[:graph:]: letters, numbers, and punctuation.
[:print:]: letters, numbers, punctuation, and whitespace.
[:space:]: space characters (basically equivalent to \s).
[:blank:]: space and tab.


```{r}
# Matches movies with digits
movie_titles[str_detect(movie_titles,"\\d")]

# matches all containg digits OR letter s
movie_titles[str_detect(movie_titles,"[s\\d]")]

# matches punctuation and upperclass letters
movie_titles[str_detect(movie_titles,"[:punct:][:upper:]")]

# List all movies that contain "Grey" or "Gray"
movie_titles[str_detect(movie_titles,"Gr[ae]y")]
```


### 9. Alternation with pipe and question mark

By connecting multiple words with a pipe operator | you both match words before the pipe and after. You can chain as many words as you like. 
By using parentheses you can group certain words together.

```{r}
str_detect(c("abc", "def", "ghi"), "abc|def")
# will return: [1]  TRUE  TRUE FALSE

# parentheses are used to override the default precedence rules:
movies <- c("Karate Kid 2, Distributor: Columbia, 58 Screens","Finding Nemo, Distributors: Pixar and Disney, 10 Screens","Finding Harmony, Distributor: Unknown, 1 Screen","Finding Dory, Distributors: Pixar and Disney, 8 Screens")

# these two lines of code will return different results
str_view(movies,"Finding Nemo|Harmony|Dory")

str_view(movies,"Finding (Nemo|Harmony|Dory)")

str_match(movies,"Finding (Nemo|Harmony|Dory)")
```

Question mark ? makes the preceding group or character optional, causing a regular expression to match even if a certain part of the pattern is missing. 
```{r}
# Match both Screen and Screens by making the last "s" optional
str_view_all(movies, pattern = "Screens?")
```

### 10. Repetition

The repetition operators control how many times a pattern matches:
?: 0 or 1.
+: 1 or more.
*: 0 or more.

To specify the number of matches precisely use {}:
{n}: exactly n
{n,}: n or more
{n,m}: between n and m
abc*        matches a string that has ab followed by zero or more c
abc+        matches a string that has ab followed by one or more c
abc?        matches a string that has ab followed by zero or one c
abc{2}      matches a string that has ab followed by 2 c
abc{2,}     matches a string that has ab followed by 2 or more c
abc{2,5}    matches a string that has ab followed by 2 up to 5 c
a(bc)*      matches a string that has a followed by zero or more copies of the sequence bc
a(bc){2,5}  matches a string that has a followed by 2 up to 5 copies of the sequence bc
