---
title: "Text mining using tm, qdap, wordcloud and dendextend R packages"
author: "Illarion Jabine"
date: "03/02/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

### Required packages:

* [tm]: Comprehensive text mining package
* [qdap]: Automates many of the tasks associated with quantitative discourse analysis of transcripts containing discourse
including frequency counts of sentence types, words, sentences, turns of talk, syllables and other assorted
* [wordcloud]: word clouds, visualize differences and similarity between documents
* [plotrix]: Various plotting functions
* [dendextend]: Extends 'dendrogram' functionality in R
* [NLP]: Basic classes and methods for Natural Language Processing
* [dplyr]


### Key terms
 * Bag of words
 * Corpus
 * Document-term matrix (DTM)/ term-document matrix(TDM)
 * Text pre-processing
 * Word cloud
 * Visualizing common words
 * Pyramid plot
 * Word networks
 * Word clustering
 * Distance matrix
 * Word dendrogram
 
## Introduction

Text mining can be defined as a process of extracting actionable insights from text.
The text mining workflow starts with the problem definition which is derived from gaining subject matter expertise.
Then different raw text sources are identified. The source text is stil in unorganized state, i.e. in semi and un-structured.
The text should go pre-processing phase, it must be organized, useful features extracted from it, and only after the text is in organized state
that the analysis can begin.
There are two major approaches to text analytics:
 1. Semantic parsing
 2. Bag of words
In this manual I will use bag of words approach.
The tm text mining package work with a document corpus (corpora in plural).
Corpora are collections of documents containing (natural language) text.
A corpus can also have metadata information about the text documents.
There are two kinds of the corpus data type, the permanent corpus - PCorpus, and the volatile corpus - VCorpus.
The volatile corpus is held in your computer's RAM rather than saved to disk.
Once the raw text data is stored in corpus, the tm package can convert it into DTM or TDM matrix.
DTM is constructed in the following way:
 * each row represents one document,
 * each column represents one term,
 * each cell contains the frequency of appearances of that term in the document.
TDM is a transposed version of DTM.

### Useful Links
 * For a comprehensive list of R NLP packages see <https://cran.r-project.org/web/views/NaturalLanguageProcessing.html>.


### Load the libraries
Let's first load the libraries. 
```{r loading packages, message=FALSE, warning=FALSE}
library(tm)
library(qdap)
library(wordcloud)
library(dendextend)
library(plotrix)
library(dplyr)
```


### 1. Loading text files
We are going to work with 2 twitter csv extracts: one about coffee and the other about chardonnay 
```{r Loading text files, message=FALSE, warning=FALSE}
# important to use the argument stringsAsFactors = FALSE
coffee <- read.csv("C:/Users/Illarion/Documents/R Projects/Text Analytics New/data/text_mining_using_tm/coffee.csv", stringsAsFactors = FALSE)
chardonnay <- read.csv("C:/Users/Illarion/Documents/R Projects/Text Analytics New/data/text_mining_using_tm/chardonnay.csv", stringsAsFactors = FALSE)

# We are interested only in text columns, to isolate text from tweets:
coffee_tweets <- coffee$text
chardonnay_tweets <- chardonnay$text
```

### 2. Creating a corpus of documents
Text corpus is a large and structured set of texts or collections of documents containing (natural language) text. 
It is used  to do statistical analysis and hypothesis testing, checking occurrences or validating linguistic rules.
A corpus is the main structure for managing documents in tm (text mining) package. 
VCorpus() function creates a volatile corpus object with conent of each document and additinal metadata.
Metadata is used to annotate text documents or whole corpora with additional information.
Use meta() function to update metadate.

```{r creating corpus, message=FALSE, warning=FALSE}
# To make a volatile corpus, R needs to interpret each element in our vector of text, coffee_tweets and chardonnay_tweets, as a document.
# There are two source function called VectorSource() and DataframeSource() to create source objects for VCorpus() function. In our case I will use VectorSource().
#The data frame passed to DataframeSource() must have a specific structure:
# Column one must be called doc_id and contain a unique string for each row.
# Column two must be called text with "UTF-8" encoding.
# Any other columns, 3+ are considered metadata and will be retained as such.

# Make a vector source from coffee_tweets
coffee_source <- VectorSource(coffee_tweets)
chardonnay_source <- VectorSource(chardonnay_tweets)

# Now we can create a volatile corpus from the vector source
coffee_corpus <- VCorpus(coffee_source )
chardonnay_corpus <- VCorpus(chardonnay_source )

# Individual elements of corpus can be accessed using [[]]
chardonnay_corpus[[20]]
# Print the contents of the 10th tweet in chardonnay_corpus
content(chardonnay_corpus[[20]]) # or
chardonnay_corpus[[20]][1]

# To access or modify the metadata:
meta(chardonnay_corpus[[20]])
meta(chardonnay_corpus[[20]], tag = "author") <- "Bob Brown"
```

### 3. Corpus text pre-processing 
tm package offer a comprehensive tm_map function that allows to apply transformation functions (also denoted as mappings) to corpora.
tm_map() takes two arguments, a corpus and a cleaning function.

Common preprocessing functions include:
tolower(): Make all characters lowercase
removePunctuation(): Remove all punctuation marks
removeNumbers(): Remove numbers
stripWhitespace(): Remove excess whitespace
removeWords(): remove words from text (normally stop words)

qdap package offers other text cleaning functions:
* bracketX(): Remove all text within brackets (e.g. "It's (so) cool" becomes "It's cool")
* replace_number(): Replace numbers with their word equivalents (e.g. "2" becomes "two")
* replace_abbreviation(): Replace abbreviations with their full text equivalents (e.g. "Sr" becomes "Senior")
* replace_contraction(): Convert contractions back to their base words (e.g. "shouldn't" becomes "should not")
* replace_symbol() Replace common symbols with their word equivalents (e.g. "$" becomes "dollar")
 ! Note: For compatibility, base R and qdap functions need to be wrapped in content_transformer()
```{r  corpus preprocessing: clean_corpus function, message=FALSE, warning=FALSE}
# To make the cleaning steps easier let's write a customized function:
clean_corpus <- function(corpus){
# create a vector of stop words by combining standard stopwords with domain specific.
# As we textmine for coffee and chardonnay these words should be removed, because
# otherwise they will dominate other words, and make finding useful patterns difficult.
# To review standard stop words call stopwords("en")
  custom_stop_words <- c(stopwords("en"), "coffee", "mug","chardonnay")
  corpus <- tm_map(corpus,content_transformer(replace_abbreviation))
  corpus <- tm_map(corpus, removePunctuation)
  corpus <- tm_map(corpus, removeNumbers)
  #Convert to lower case, tolower() is a base R function
  corpus <- tm_map(corpus, content_transformer(tolower)) 
  corpus <- tm_map(corpus, removeWords, custom_stop_words)
  # Eliminating extra whitespace
  corpus <- tm_map(corpus, stripWhitespace)
  return(corpus)
}

# Now we have to apply the pre-processing tranformations to our corpus of documents:
coffee_corpus_clean <- clean_corpus(coffee_corpus)
chardonnay_corpus_clean <- clean_corpus(chardonnay_corpus)

```

 3.1 Stemming
Stemming is another useful preprocessing step. 
It involves word stemming itself, which is a reduction of words to a root form and stem completion.
stemDocument() function from tm package allows you to get to a word's root.
stemCompletion() re-completes the words using completion dictionary as the reference corpus.
 
### 4. Document Term/Term Document Matrix
A common approach in text mining is to create a term-document matrix from a corpus. 
Term-document matrix is the basis for bag of words text mining.
TermDocumentMatrix() and DocumentTermMatrix() (depending on whether you want terms as rows and
documents as columns, or vice versa) are used to create DTM/TDM.
```{r Document Term/Term Document Matrix, message=FALSE, warning=FALSE}

# creat DTM (documents in rows)
coffee_dtm <- DocumentTermMatrix(coffee_corpus_clean)
chardonnay_dtm <- DocumentTermMatrix(chardonnay_corpus_clean)

# Convert DTMs to a matrix
coffee_m <- as.matrix(coffee_dtm)
chardonnay_m <- as.matrix(chardonnay_dtm)
dim(coffee_m)
# Review a portion of the matrix, take first 30 rows with terms "milk" and "black"
coffee_m[1:30, c("milk","black")]

# create TDM (terms in rows)
coffee_tdm <- TermDocumentMatrix(coffee_corpus_clean)
chardonnay_tdm <- TermDocumentMatrix(chardonnay_corpus_clean)

# Convert TDMs to a matrix
coffee_m_tdm <- as.matrix(coffee_tdm)
chardonnay_m_tdm <- as.matrix(chardonnay_tdm)

# Note coffee_m_tdm is a transposed version of coffee_m
dim(coffee_m_tdm)

coffee_m_tdm[c("milk","black"),1:30]
```

### 5. Frequent Terms

Using coffee_m_tdm matrix (terms in rows, documents in columns) it is easy to calculate various statistics.
Let's find the frequent terms using base R functions
```{r frequent terms, message=FALSE, warning=FALSE}
# Calculate the row sums of coffee_m_tdm
coffee_term_frequency <- rowSums(coffee_m_tdm)

# Sort coffee_term_frequency in decreasing order
term_frequency <- sort(coffee_term_frequency, decreasing = TRUE)

# Plot a barchart of the 10 most common words
barplot(term_frequency[1:20], col = "blue", las = 2)

```

There is a fast way to get frequent terms without going through preprocessing steps.
freq_terms() from qdap package. The function accepts a text variable:
```{r frequent terms using freq_terms(), message=FALSE, warning=FALSE}
# define a custom stopword dictionary
custom_stop_words <- c(stopwords("english"), "coffee", "mug","chardonnay")

term_frequency_qdap <- freq_terms(
  coffee_tweets, # text source  
  top = 20, # specify the top number of terms to show
  at.least = 3, # the minimum character length of a word to be included
  stopwords = custom_stop_words #a vector of stop words to remove
)

# Make a frequency barchart
plot(term_frequency_qdap)
```

### 6. Word Cloud

A word cloud is a visualization of terms. 
In a word cloud, size is often scaled to frequency and the colors may indicate another measurement.
To create a word cloud I will use wordcloud() function from wordcloud package.

```{r word cloud for coffee terms, message=FALSE, warning=FALSE }
# for word cloud we need term frequencies and term names 
term_names <- names(coffee_term_frequency)

# We can construct the cloud using three colors.
# That will divide the term frequency into "low", "medium" and "high"
wordcloud(term_names, 
          coffee_term_frequency, 
          max.words = 50, 
          colors = c("grey80", "darkgoldenrod1","tomato"))

```

### 7. Visualize common words
If I want to visualize common words across multiple documents, for example
For example, for the marketing campaign how words in common can help understand similar brands features, etc.
To visualize common words used in coffee and chardonnay tweets we can use commonality.cloud() function from wordcloud package.
However, to do that, we need to create a new corpus with only two documents:
one combining all tweets about coffee and the other combining all tweets about chardonnay.
Remember our corpus "coffee_corpus_clean" and "chardonnay_corpus_clean" are composed of many individual tweets (1000 tweets). 
We need to combine them into one document per subject.
That can be done like that:
 1. paste() together all the tweets in each corpus along with the parameter collapse = " "
This collapses all tweets (separated by a space) into a single vector.
 2. Finally create a single vector containing the two collapsed documents.

```{r Visualize common words}
# Create a char vector with two elements. 
coffee_chardonnay <- c(paste(coffee_tweets,collapse = " "),
                       paste(chardonnay_tweets,collapse = " "))

# Convert to a vector source
coffee_chardonnay_source <- VectorSource(coffee_chardonnay)

# Create coffee_chardonnay_corpus
coffee_chardonnay_corpus <- VCorpus(coffee_chardonnay_source)
# As a result we have a corpus with two documents: one about coffee and the other about chardonnay.
# Now let's clean this new corpus
# Clean the corpus
coffee_chardonnay_corpus <- clean_corpus(coffee_chardonnay_corpus)

# Create coffee_chardonnay_tdm
coffee_chardonnay_tdm <- TermDocumentMatrix(coffee_chardonnay_corpus)

# Give the columns distinct names
colnames(coffee_chardonnay_tdm) <- c("Coffee","Chardonnay")

# Convert the TDM object to a matrix using as.matrix() for use in commonality.cloud() 
coffee_chardonnay_m <- as.matrix(coffee_chardonnay_tdm)

# Print a commonality cloud
commonality.cloud(coffee_chardonnay_m, max.words = 100, colors = "steelblue1")
```

### 8. Comparing the frequencies of words

To plot a cloud comparing the frequencies of words across documents.
To do that use comparison.cloud() and the steps are similar to explained in "7. Visualize common words".
I just re-use "coffee_chardonnay_m" matrix from "7. Visualize common words"
```{r Visualize dissimilar words}

comparison.cloud(coffee_chardonnay_m, max.words = 50, colors = c("orange", "blue"))

```

### 9. Pyramid plot: Polarized tag cloud

In addition to words shared across documents, we might be interested in words
that appear more commonly in one document compared to another.
That can be done with a pyramid (opposed horizontal bar) plot produced by pyramid.plot() from the plotrix package.
Some data manipulation is required to get a pyramid plot:
 1. Convert TDM matrix to a data frame with 3 columns:
  1. The words contained in each document.
  2. The counts of those words from document 1.
  3. The counts of those words from document 2.
Then execute this: 
pyramid.plot(word_count_data$count1, word_count_data$count2, word_count_data$word)
```{r Pyramid plot}
top30_df <- coffee_chardonnay_m %>%
  # Convert to data frame
  as_data_frame(rownames = "word") %>% 
  # Keep rows where word appears everywhere
  filter_all(all_vars(. > 0)) %>% 
  # Get difference in counts
  mutate(difference = Chardonnay - Coffee) %>% 
  # Keep rows with biggest difference
  top_n(30, wt = difference) %>% 
  # Arrange by descending difference
  arrange(desc(difference))


  pyramid.plot(
  # Coffee counts
  top30_df$Coffee, 
  # Chardonnay counts
  top30_df$Chardonnay, 
  # Words
  labels = top30_df$word, 
  top.labels = c("Coffee", "Words", "Chardonnay"), 
  main = "Words in Common", 
  unit = NULL,
  gap = 10,
)
```

### 10. Word network

Word networks is another way to view word connections. 
Such network shows term association and cohesion (<https://en.wikipedia.org/wiki/Cohesion_(linguistics)>).
Nodes represent individual terms, and the lines connecting the nodes are called edges and represent the connections between the terms.
Below is a word network for words associated with "java":
```{r}
word_associate(coffee_tweets, match.string = "java", 
               stopwords = c(Top200Words), 
               network.plot = TRUE, cloud.colors = c("gray85", "darkred"))
title(main = "Java Coffee Tweet Associations")
```

